DATA DRIVEN MODELS FOR APOGEE M DWARFS

### 2018-03-30 paper/draft1/draft1.tex
acknowlegments

\subsection{M dwarf Challenge}
\begin{enumerate}
\item[-] Various attempts / samples for which people have tried to make substantial catalogs of M-dwarfs with physical labels on them (\citealt{West:2011} is the largest M dwarf catalog of $\sim$ 70k sources; \citealt{Deacon:2014} -- \color{red}{Need citations!}\color{black})

@@

The APOGEE survey \citep{Majewski:2015} of the Sloan Digital Sky Survey IV (SDSS-IV) is a high resolution (R$\sim$22,500), H-band (1.5-1.7$\mu$m) survey which has observed over 260,000 stellar spectra since its fourteenth data release (DR14; \citealt{Abolfathi:2017}). Fundamental parameters for each of these stars are estimated by the APOGEE Stellar Parameter and Chemical Abundances Pipeline (ASPCAP; \citealt{Perez:2016}), which employs a $\chi^2$ fitting procedure using the FERRE code to fit ATLAS9 radiative transfer models \citep{Castelli:2004} to determine atmospheric parameters, 15 chemical abundances and micro-turbulence parameters (T$_{\rm eff}$, log g, [Fe/H], [α/M], [C/M], [N/M]; \citealt{Meszaros:2012}).

@@

The spectral type sample consists of 51 sources, spanning M0-M9 (Figure \ref{fig:train_dist}) cross-matched from the \citealt{West:2011} catalog of 78,841 M dwarfs from SDSS. For each source in the catalog, spectral types were determined both through an automated routine using The Hammer \citep{Covey:2007} and by visual inspection.

The physical parameter sample consists of 30 sources distributed over $3072 < T_{\rm eff} < 4131K$, and -0.48 $<$ [Fe/H] $<$ 0.49 (Figure \ref{fig:train_dist}) drawn from \citealt{Mann:2015}.
\color{red}{why logg not included as atmospheric parameter?}\color{black}

@@

\subsection{Data Processing}
Describe ASPCAP spectra files and Ness continuum normalization procedure.

@@

\item[-] Two tests of consistency were used to assess the validity of our derived test labels: Figure \ref{fig:west_selftest} displays the label self-test for the spectral type model, and Figure 2 displays 
\end{enumerate}

@@

>>>>

\subsection{M dwarf Challenge}
\begin{enumerate}
\item[-] Various attempts / samples for which people have tried to make substantial catalogs of M-dwarfs with physical labels on them (\citealt{West:2011} is the largest M dwarf catalog of $\sim$ 70k sources...)

@@

The APOGEE survey \citep{Majewski:2015} of the Sloan Digital Sky Survey III (SDSS-III; \citealt{Eisenstein:2011}) is a high resolution (R$\sim$22,500), H-band (1.5-1.7$\mu$m) survey which has observed over 260,000 stellar spectra since its fourteenth data release (DR14; \citealt{Abolfathi:2017}). Fundamental parameters for each of these stars are estimated by the APOGEE Stellar Parameter and Chemical Abundances Pipeline (ASPCAP; \citealt{Perez:2016}), which employs a $\chi^2$ fitting procedure using the FERRE code to fit ATLAS9 radiative transfer models \citep{Castelli:2004} to determine atmospheric parameters, 15 chemical abundances and micro-turbulence parameters (T$_{\rm eff}$, log g, [Fe/H], [$\alpha$/M], [C/M], [N/M]; \citealt{Meszaros:2012}).

@@

The spectral type sample consists of 51 sources, spanning M0-M9 (Figure \ref{fig:train_dist}) cross-matched from the \citealt{West:2011} (hereafter W11) catalog of 78,841 M dwarfs from SDSS. For each source in the catalog, spectral types were determined both through an automated routine using The Hammer \citep{Covey:2007} and by visual inspection to a reported accuracy of $\pm$1 type.

The physical parameter sample consists of 30 sources with reference labels distributed over $3072 < T_{\rm eff} < 4131K$, and -0.48 $<$ [Fe/H] $<$ 0.49 (Figure \ref{fig:train_dist}) drawn from \citealt{Mann:2015} (hereafter M15). The M15 catalog in total contains 183 sources selected from CONCH-SHELL \citep{Gaidos:2013} and with infrared spectra taken in follow up from the SpeX Spectograph. Effective temperatures are measured by comparison of BT-SETTL atmospheric models \citep{Allard:2011} to optical spectra. A subsample of 29 sources with measured angular diameters and bolometric fluxes from long-baseline optical interferometry are used to callibrate effective temperature measurements. Relative uncertainty from the callibration sample is measured to be \color{red}{$\sim$30K (check this)}\color{black}, which added in quadrature to the scatter from model comparison gives a total reported uncertainty of $\sim$60K in T$_{\rm eff}$.

Spectral types reported by M15 for the sources using optical molecular band indices (by the method of \citealt{Lepine:2013}) and ranged K7.3--M4.9 for the subsample we used, although these labels were not included in the model training.

\color{red}{why logg not included as atmospheric parameter?}\color{black}

@@

\subsection{Data Processing}

APOGEE data products come in several types of files: ap1D, apVisit, apStar, and aspcapStar. SNR of stars in our sample.

Describe ASPCAP spectra files and Ness continuum normalization procedure.

@@

\item[-] Two tests of consistency were used to assess the validity of our derived test labels: Figure \ref{fig:west_selftest} displays the label self-test for the spectral type model, and Figure 2 displays

\item[-] Vsini and logg were not included as training parameters. Vsini is less import of a parameter as the sample consists of mainly old, slowly rotating sources (check what are ages?). Logg because?
\end{enumerate}

@@

%%==========================================================================================
\acknowledgements
We would like to acknowledge Derek Homeier (Universit{\"a}t Heidelberg), Wolfgang Bradner (MPIA), Hans Walter Rix (MPIA), Anna-Christina Eilers (MPIA), Anna Ho (Caltech), and Jackie Faherty (AMNH) for constructive discussions in the process of this project, as well as Bertrand Goldman (MPIA) and those who have supported the internship program at the Max Planck Institute f{\"u}r Astronomie for providing JB with funding and hospitality. Software: The Cannon, numpy, matplotlib...

\color{red}{HOGG: SDSS acknowledgments... Grants and stuff...}\color{black}
